---
title: "Do the rich get richer in Basel?"
date: '2022-06-23'
author: "The R Bootcamp"
output: 
  html_document: 
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(tidyverse)
require(patchwork)

basel = read_csv("../1_Data/taxation.csv")
```


:::{.column-body}
[column-body]{style="color:red"} 
Inequalities in income and wealth are a growing social issue. This analysis seeks to quantify inequality in Basel, Switzerland, by analyzing the widening of wealth and income gaps between the city's quarters.    
:::

:::{.column-body-outset}
[column-body-outset]{style="color:red"} 
Inequalities in income and wealth are a growing social issue. This analysis seeks to quantify inequality in Basel, Switzerland, by analyzing the widening of wealth and income gaps between the city's quarters.    
:::

:::{.column-page}
[column-page]{style="color:red"} 
Inequalities in income and wealth are a growing social issue. This analysis seeks to quantify inequality in Basel, Switzerland, by analyzing the widening of wealth and income gaps between the city's quarters.    
:::

:::{.column-margin}
[column-margin]{style="color:red"} 
Inequalities in income and wealth are a growing social issue. This analysis seeks to quantify inequality in Basel, Switzerland, by analyzing the widening of wealth and income gaps between the city's quarters.    
:::

```{r}
#| column: margin
#| fig-asp: 0.7
#| echo: false
# create quarter factor
basel <- basel %>%
  arrange(year, income_mean) %>% 
  mutate(quarter = as_factor(quarter)) 
  
# create income plot
plot_income <- basel %>% 
  ggplot(aes(x = year, y = income_mean, col = quarter)) +
  labs(x = "Year", y = "Income")

# create income plot
plot_wealth <- basel %>%
  ggplot(aes(x = year, y = wealth_median, col = quarter)) +
  labs(x = "Year", y = "Wealth")

# create joint plot
plot_income + plot_wealth  + 
  plot_layout(guides = "collect") +
  plot_annotation(title = "Inequality in Basel",
                  subtitle = "Income and wealth development from 2001 to 2017",
                  caption = "Source: Open Data Basel",
                  tag_levels = "A") & 
  geom_point() & 
  geom_line() &
  scale_color_viridis_d() &
  theme_minimal() &
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.height = unit(.01,"in"))
  
```